Andres Karjus

### 2020-12-01 index.Rmd
postdoc updates

<div style="text-align: justify; padding-left:1px"> 
I am a research fellow at the [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, studying language and cultural dynamics. I recently defended my PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/) on lexical competition and communicative needs in language, supervised by [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/).
I am also engaged in teaching stats and data visualization for the humanities and social sciences (see the tab above). On the right is the picture obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Competition, selection and communicative need in language change. PhD [thesis](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf), University of Edinburgh
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [preprint](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. [Language Dynamics and Change](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

- I am engaged in teaching stats and R to [Edinburgh Uni psychology](https://www.ed.ac.uk/ppls/psychology) masters students (since 2017).
- [I am available for booking](http://writingcentre.ppls.ed.ac.uk/appointments/) year around as a written communication consultant via the Edinburgh University School of Psychology, Philosophy and Language Sciences Writing Centre - specializing in writing about and presenting (graphs and figures of) data and data analysis results. This service is available to students of the School only. To book, go to the [appointments](http://writingcentre.ppls.ed.ac.uk/appointments/) page > Book an appointment > look for my name in the list of Linguistics tutors.

@@

- Developed and co-taught a course on data analysis for digital humanities at the University of Tartu in the spring of 2016.
- Invited lecturer, corpus linguistics using R course for the [Academia Salensis](http://academiasalensis.org) summer school of 2015.
- Assistant for courses on language technology and artificial intelligence, Department of Computer Science, University of Tartu, 2014-2016.
<br><br>

Feel free to get in touch if you are interested in talking about organizing a workshop on anything related to data science and statistics for the humanties, data visualization, R, corpus linguistics, natural language processing applications, digital humanities, etc.
</div>
<br>

## Other things
### Semi-academic
<div style="text-align: justify">

- I've been helping my supervisor Prof. Kenny Smith carry out occasional language evolution workshops in primary schools around Edinburgh and Lothian, in 2018/2019 and 2019/2020.
- Entered [a video](https://www.youtube.com/watch?v=t8InfFEIINc) to Science Magazine's 11th [Dance Your Ph.D contest](https://www.sciencemag.org/projects/dance-your-phd). The idea is to explain your research topic to the general public though interpretive dance. We didn't win but it was a fun one to do!
- Co-organized the bimonthly [Edinburgh Language Lunch](http://groups.inf.ed.ac.uk/langlunch/) during 2016-2018. 
- Co-organized the [Linguistics and English Language Postgraduate Conference](http://www.lel.ed.ac.uk/~pgc/) of 2017 in Edinburgh, the [Why Linguistics Conference](http://whylinguistics.ut.ee) (2015) and the [Graduate Conference on Areal Linguistics, Grammar and Contacts](http://arealling.ut.ee) (2012) in Tartu, Estonia.

### Non-academic

Besides research and teaching and consulting and whatnot, I (fortunately) also do other things, which currently mostly consist of dance (lindy hop, salsa, bachata), jazz, outdoorsy stuff (cycling, hillwalking), boardgames, and various other things.

### Before Scotland

Before coming to Edinburgh, I worked as a teaching assistant in informatics at the University of Tartu (2015-2016), before that studied artificial intelligence and natural language processing at KU Leuven (MSc) and linguistics at the University of Tartu (BA, MA). I was also affiliated 2016-2019 as a (part-time) junior researcher with the University of Tartu [EKKAM](http://ekkam.ut.ee/en/) sociolinguistics group, doing data analysis and agent-based models. 
During my pre-PhD studies I also went on exchanges to the University of Iceland and the University Vienna, attended a dozen-odd academic summer schools, taught Icelandic to art students and Estonian to Norwegian teachers, worked as an assistant at the Estonian Wordnet project, and did internships at CrossLang NV in Belgium and at the (then still existing linguistics department of the) Max Planck Institute for Evolutionary Anthropology in Leipzig. In earlier years, I worked various studenty sort of jobs to support my studies (not to list them all, but one of them involved swords and another one regularly mowing grass off the roof of a small yellow building).
</div>
<br>

@@

<div style="text-align: center; align:center"> 
<IMG SRC="round.png"> 
<br>
<br>
<br>
<a href="mailto:a.karjus--at--sms.ed.ac.uk">&nbsp; email</a>  |  <a href="https://twitter.com/AndresKarjus">twitter</a>
<br>
</div>

>>>>

<div style="text-align: justify; padding-left:1px"> 
I am a senior research fellow at the [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, working on **language and culture dynamics** using large corpora, machine learning, but also human experiments. I recently defended my PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/) on lexical competition and communicative needs in language, supervised by [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/).
I am also engaged in teaching **stats and data visualization** for the humanities and social sciences (see the tab above). On the right is the (of course slightly outdated) photo obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [journal link](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

- My postdoc does not include teaching activities, but I do occasional guest lectures; most recently for the Data Science and Digital Humanities programme at the University of Tartu.

@@

- I was engaged in teaching stats and R to [Edinburgh Uni psychology](https://www.ed.ac.uk/ppls/psychology) masters students 2017-2019.
- And worked for the Edinburgh University School of Psychology, Philosophy and Language Sciences Writing Centre 2017-2020 as awritten communication consultant, specializing in writing about and presenting data and data analysis results.
- Developed and co-taught a course on data analysis for digital humanities at the University of Tartu in the spring of 2016.
- Lectured on corpus linguistics and R for the [Academia Salensis](http://academiasalensis.org) summer school of 2015.
- Worked as a teaching assistant for courses on language technology and artificial intelligence, Department of Computer Science, University of Tartu, 2014-2016.
<br><br>

Feel free to get in touch if you are interested in talking about organizing a workshop on anything related to data science and statistics, data visualization, R, corpus linguistics, natural language processing applications, digital humanities, etc.
</div>
<br>

## Other things
### Semi-academic & science popularization stuff
<div style="text-align: justify">

- Hosted a student via the job shadowing (töövari) programme of the Estonian Tagasi Kooli programme (November 2020).
- I helped my supervisor Prof. Kenny Smith carry out a number of linguistics workshops in primary schools around Edinburgh and Lothian, in school years 2018/2019 and 2019/2020.
- Entered [a video](https://www.youtube.com/watch?v=t8InfFEIINc) to Science Magazine's 11th [Dance Your Ph.D contest](https://www.sciencemag.org/projects/dance-your-phd). The idea is to explain your research topic to the general public though interpretive dance. We didn't win but it was a fun one to do!
- Co-organized the bimonthly [Edinburgh Language Lunch](http://groups.inf.ed.ac.uk/langlunch/) during 2016-2018. 
- Co-organized the [Linguistics and English Language Postgraduate Conference](http://www.lel.ed.ac.uk/~pgc/) of 2017 in Edinburgh, the [Why Linguistics Conference](http://whylinguistics.ut.ee) (2015) and the [Graduate Conference on Areal Linguistics, Grammar and Contacts](http://arealling.ut.ee) (2012) in Tartu, Estonia.

### Non-academic stuff

Besides research and teaching and consulting and whatnot, I (fortunately) also do other things, which mostly consist of dance (lindy hop, salsa, bachata), jazz, outdoorsy stuff (cycling, hillwalking), boardgames, and various other things.

### Before Edinburgh

In the more distant past, I worked as a teaching assistant in informatics at the University of Tartu (2015-2016), before that studied artificial intelligence and natural language processing at KU Leuven (MSc) and linguistics at the University of Tartu (BA, MA). I was also affiliated 2016-2019 as a (part-time) junior researcher with the University of Tartu [EKKAM](http://ekkam.ut.ee/en/) sociolinguistics group, doing data analysis and agent-based models. 
During my pre-PhD studies I also went on exchanges to the University of Iceland and the University Vienna, attended a dozen-odd academic summer schools, taught Icelandic to art students and Estonian to Norwegian teachers, worked as an assistant at the Estonian Wordnet project, and did internships at CrossLang NV in Belgium and at (the old) Linguistics Departent of the Max Planck Institute for Evolutionary Anthropology in Leipzig. In earlier years, I worked various studenty sort of jobs to support my studies (for a seller of swords, for a seller of cars, for a minder of horses and tourists).
</div>
<br>

@@

<div style="text-align: center; align:center"> 
<IMG SRC="round.jpg"> 
<br>
<br>
<br>
&nbsp;&nbsp;&nbsp;<a href="mailto:andres.karjus--at--tlu.ee">email</a>  |  <a href="https://twitter.com/AndresKarjus">twitter</a>
<br>
</div>
